---
title: "Post-Disaster Demographic Changes"
author: "Matt Khinda, Jeb Polstein, Slide Kelly"
date: "10/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Research Question  

How well do demographic factors predict post-hurricane Sandy displacement? Which factors are most strongly correlated with displacement?  

##Prior Research

Faber (2015) finds that flood risk in New York City during Hurricane Sandy varied by race and poverty. Black people were more likely than those of other races to live in areas that flooded, Latinx people were less likely to live in areas that flooded, and poverty rates were higher in areas that flooded than those that did not. Fussell and Harris (2014), studying displacement in New Orleans after Hurricane Katrina, find that renters were more likely to be displaced than homeowners. 


##Data Sources

a

##Load Data
```{r load packages}
library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
library(ggplot2)
```

```{r explore ACS data, include=FALSE}
v19 <- load_variables(2019, "acs5", cache = TRUE)
View(v19)
```

```{r load ACS data}


```

##References

Faber, Jacob William. “Superstorm Sandy and the Demographics of Flood Risk in New York City.” *Human Ecology* 43, no. 3 (2015): 363–78. http://www.jstor.org/stable/24762762.

Fussell, Elizabeth, and Elizabeth Harris. “Homeownership and Housing Displacement after Hurricane Katrina among Low-Income African-American Mothers in New Orleans.” *Social Science Quarterly* 95, no. 4 (2014): 1086-1100. doi:10.1111/ssqu.12114. 

United States Census Bureau. American Community Survey, 5-year estimates. 2011-2016.

